### 2024-07-24 reports/paper/main.tex
paper: Initial Arxiv template

% Arxiv styling
% Based on https://github.com/kourgeorge/arxiv-style
\documentclass{article}
\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{floatrow}
%\usepackage{natbib}
\usepackage{doi}
\usepackage{graphicx}
\usepackage{subcaption}

% Document
\begin{document}

%% Title
\title{PRELIMINARY. Leaf compression for tiny Random Forests}

%% Author info
\author{
    Jon Nordby \\
	Soundsensing AS \\
	\texttt{jon@soundsensing.no} \\	
}

% Remove the date
\date{}

\maketitle
\renewcommand{\abstractname}{\vspace{-\baselineskip}} % erase the space between authors and abstract

%% Abstract
\begin{abstract}	\noindent

% Keywords. Maximum 5
\noindent \textbf{Keywords}:

\end{abstract}

\section{Introduction}

\noindent

\newpage
\section{Background}

\subsection{Related work}

\section{Methods}

\newpage
\section{Results}

%% DISCUSSION
\section{Discussion}

\subsection{Further work}

%\newpage
\section{Conclusions}

\section*{Acknowledgements}
\noindent

% References
\newpage
\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}